---
title: "Structure test"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Structure test}
  %\usepackage[utf8]{inputenc}
---

In this vignette we show that the Mantel test will be significant even if there is no structure.

We start by generating a dated tree with root in 2000 and small $N_e g=1$ year relative to sampling frame of 10 years:

```{r}
library(BactDating)
library(ape)
set.seed(0)
dates=runif(100,0,10)
dt=simcoaltree(dates,alpha=1)
dates=dates+2000-dt$root.time
dt$root.time=2000
plot(dt,show.tip.label = F)
axisPhylo(backward = F)
```

Next we generate a phylogenetic tree using a Poisson(10) mutation model:
```{r}
tree=simobsphy(dt,mu=10,model='poisson')
plot(tree,show.tip.label = F)
axisPhylo(1,backward=F)
```

A root-to-tip regression analysis works perfectly:

```{r}
res=roottotip(tree,dates)
```

The bactdate analysis also works perfectly:

```{r}
r=bactdate(tree,dates,model='poisson',updateRoot = F)
plot(r,'trace')
plot(r,'treeCI',show.tip.label=F)
print(r)
```

There is strong association between genetic and temporal distances, even if this is not actually an issue for dating:

```{r}
tim=as.matrix(dist(dates))
gen=cophenetic.phylo(tree)
plot(tim[upper.tri(tim)],gen[upper.tri(gen)],xlab='Time distance',ylab='Genetic distance')
```

Mantel test from the vegan package:
```{r}
vegan::mantel(gen,tim)
```

Mantel test from the ape package:
```{r}
ape::mantel.test(gen,tim)
```

Mantel test from the ade4 package:
```{r}
ade4::mantel.rtest(as.dist(gen),as.dist(tim))
```