---
title: "Pseudo-residual analysis for relaxed clock"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  html_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Pseudo-residual analysis for relaxed clock}
  %\usepackage[utf8]{inputenc}
---

We start by generating a dated tree with no structure and a small effective population size 
$N_e g=1$ year relative to the sampling frame of 10 years between 2010 and 2020:

```{r}
library(BactDating)
library(ape)
library(ValidateDating)
set.seed(1)
dates=runif(100,2010,2020)
dt=simcoaltree(dates,alpha=1)
plot(dt,show.tip.label = F)
axisPhylo(backward = F)
```

Next we generate a phylogenetic tree using a ARC(10,5) mutation model:
```{r}
tree=simobsphy(dt,mu=10,sigma=5,model='arc')
plot(tree,show.tip.label = F)
axisPhylo(1,backward=F)
```

A root-to-tip regression analysis works not so well:

```{r}
res=roottotip(tree,dates)
```

Run analysis with incorrect model:
```{r}
r=runDating(tree,dates,algo='BactDating',model='poisson')
print(r)
```

Pseudo-residual analysis:
```{r}
plotProbBranches(r)
plotResid(r)
print(testResid(r))
```

Run analysis with correct model:
```{r}
r=runDating(tree,dates,algo='BactDating',model='arc')
print(r)
```

Pseudo-residual analysis:
```{r}
plotProbBranches(r)
plotResid(r)
print(testResid(r))
```
